MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-03-13 MAESTROeX_I/paper.tex
clean up alg description more

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \}
\end{eqnarray}
\end{description}

@@

The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
\end{equation}
where
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
\qquad
\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
\qquad
p_0^{\nph} = \frac{ p_0^n + p_0^{n+1,\star} }{2},
\end{equation}

@@

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2)}
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2)}
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
\end{eqnarray}
\end{description}
where $p_0^\nph$ is defined as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.

@@

\section{Old Notes}

\section{Computing $w_0$ For Spherical Problems}
Recall that we want to solve
\begin{equation}
\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
\end{equation}
First, note that $\partial p_0/\partial r = -\rho_0 g$.
The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
Let's discretize this as follows:
\begin{equation}
\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
\end{equation}
Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
\begin{equation}
\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
\end{equation}
As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
The interface to the routine in the algorithm description is
{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$

After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.

@@

>>>>

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
\end{eqnarray}
\end{description}
where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.

@@

The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
\end{equation}
where
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
\qquad
\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
\end{equation}

@@

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2)}
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2)}
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
\end{eqnarray}
\end{description}
where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.

@@

\section{Old Notes (not for inclusion in paper)}

\subsection{$\eta_\rho$ Notes}
After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.

@@

\subsection{Computing $w_0$ For Spherical Problems}
Recall that we want to solve
\begin{equation}
\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
\end{equation}
First, note that $\partial p_0/\partial r = -\rho_0 g$.
The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
Let's discretize this as follows:
\begin{equation}
\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
\end{equation}
Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
\begin{equation}
\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
\end{equation}
As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
The interface to the routine in the algorithm description is
{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
